---
title: "Myanmar Vulnerability"
author: "Eva Constantaras"
date: "September 4, 2018"
output: 
  html_document:
     theme: cerulean
     highlight: pygments 
     toc: true
     toc_float: true
---

```{r setup, include=FALSE, warning=F, message=F}
knitr::opts_chunk$set(echo = TRUE)
```

## Load packages 


```{r loading, warning=F, message=F}
library(ggplot2)
library(tidyverse)
library(dplyr)
library(sf)
library(readr)
library(DT)
```

##Import geographic data

Data from http://geonode.themimu.info/layers/geonode%3Amyanmar_township_boundaries 

```{r geographic, echo=TRUE, warning=F, message=F}
myanmar_ts <- "data/myanmar_township_boundaries.shp"

townships<-st_read(myanmar_ts)

ggplot(townships)+
  geom_sf()
```

##Import vulnerability data
The data is from http://themimu.info/vulnerability-in-myanmar Note that no codebook is available and the data is sourced from several different agencies. I reformatted the data in Excel before uploading to R. 

```{r vulnerability, echo=TRUE, warning=F, message=F}
library(readxl)

vuln <- read_excel("data/Myanmar_MIMU_Vulnerability.xls")

str(vuln)

vuln<-subset(vuln, select=-c(`State/Region Pcode`, `State/Region Name`, `District Pcode`, `District Name`, `Township Name`, `Number of Village Tracts`, `Number of Wards`))

```
##Clean data
```{r clean, echo=TRUE, warning=F, message=F}
#gather to create a column of Indicators
vuln_gather <- gather(data=vuln, key="Indicator", value = "Measure", -`Township Pcode`, na.rm=F)

#join datasets
myanmar <- left_join(vuln_gather, townships, by=c(`Township Pcode`="TS_PCODE"))

#narrow down indicators 
myanmar_dev <- myanmar %>% 
  filter(myanmar$Indicator %in% c("Avg (safe sanitation + improved drinking water)", 
                                  "Average good roof and wall",
                                  "Electricity for Lighting %__1" ,
                                  "Literacy: % literate"))


#rename indicators for clarity
myanmar_map_set <- mutate(myanmar_dev, map_labels=case_when(
                      Indicator =="Avg (safe sanitation + improved drinking water)"~"Clean water", 
                      Indicator =="Literacy: % literate"~"Literacy",
                      Indicator =="Average good roof and wall"~"Roof and walls",
                      Indicator == "Electricity for Lighting %__1"~"Electric lighting"))
unique(myanmar_map_set$map_labels)
```
##Map selected indicators 
```{r map, echo=TRUE, warning=F, message=F}


ggplot(myanmar_map_set)+
  geom_sf(aes(fill=Measure*100), color="white", size=.1)+
  scale_fill_distiller(direction=1, name="Population %") +
  labs(title="Uneven development across Myanmar", caption="Source: themimu.info/")+
  facet_wrap(~map_labels)+
  theme_void() +
  theme(panel.grid.major = element_line(colour = 'transparent'))
```


##Identify vulnerable states

Now that we have a detailed picture of varied development across townships, I want to look at the overall proportion of vulnerable populations by state and determine whether states with ethnic minority populations were more vulnerable since 2014. The data is from http://themimu.info/vulnerability-in-myanmar 



```{r chart data, echo=TRUE, warning=F, message=F}
#bring in state names 
myanmar_chart_2 <- vuln %>% 
                  left_join(townships, by=c(`Township Pcode`="TS_PCODE"))
```
##Proportion vulnerable by state



```
#calculate total population and total vulnerable population by state
```{r chart state data, echo=TRUE, warning=F, message=F}
myanmar_sum <- myanmar_chart_2 %>% 
  group_by(ST) %>% 
  summarize(total_pop=sum(`Total Pop Both sexes`), total_vp=sum(`Approximate Vulnerable Population`)) 
  datatable(myanmar_sum)
  

  
#Calculate percentage of population that is vulnerable
  
myanmar_sum <- mutate(myanmar_sum, perc_vuln_pop=total_vp/total_pop*100) %>% 
  arrange(desc(perc_vuln_pop))
datatable(myanmar_sum)
```
##Create chart of states with the highest proportion of vulnerable people
```{r chart 2 data, echo=TRUE, warning=F, message=F}
ggplot(data = subset (myanmar_sum, ST %in% ST [1 : 9]))+ 
       aes(x=reorder(ST, perc_vuln_pop), y=perc_vuln_pop, fill=ST)+
  geom_bar(stat="identity")+
  coord_flip()+
  xlab("States") + ylab("Vulnerable population (%)")+
  labs(title="Ethnic states most vulnerable before current crisis", 
       subtitle = "Over half of citizens vulnerable in the poorest and most marginalized ethnic states",
       caption="Source: Vulnerability Assessement, 2014 Census, themimu.info/")+
      scale_fill_brewer()+
      theme(legend.position="none")
```